\documentclass{article}
\usepackage[affil-it]{authblk}

\usepackage{graphicx}
\usepackage[space]{grffile} % extends the file name processing of package graphics

%\usepackage{latexsym} % add some characters from the "lasy" fonts
\usepackage{amsfonts,amsmath,amssymb}

\usepackage[utf8]{inputenc}

\usepackage{OI-Interface}

\hypersetup{%
  pdftitle={},
  pdfauthor={},
  colorlinks=true,
  linkcolor=DarkCyan,
  citecolor=DarkGreen,
  filecolor=DarkCyan,
  menucolor=DarkOrange,
  urlcolor=DarkCyan,
  pdfborder={0 0 0}}

\usepackage{textcomp}
\usepackage{longtable}
\usepackage{multirow,booktabs}
\usepackage[authoryear,round]{natbib}

%==============================================================================
\newcommand*{\ROW}{} % predefined here, redefined later
\newcommand*{\ROWTITLE}{} % predefined here, redefined later

\newcommand{\oops}[1]{\DarkRed{#1}}
\newcommand{\KEYWORD}[1]{\texttt{#1}} % FITS keyword
\newcommand{\STRING}[1]{\texttt{'#1'}} % FITS string value
\newcommand{\VARIABLE}[1]{\texttt{\textsl{#1}}}
%==============================================================================


\begin{document}

\title{Interface to Image Reconstruction}

\author{Éric Thiébaut$^1$, Jonathan Léger$^1$, Michel Tallon$^1$, John
  Young$^2$, Gilles Duvert$^3$, Guillaume Mella$^3$, Laurent Bourges$^3$ \&
  Martin Vannier$^4$}

\affil{$^1$CRAL, $^2$University of Cambridge, $^3$JMMC/IPAG, $^4$Lagrange}


\date{\today}

\maketitle

%==============================================================================
\section{Introduction}

Many image reconstruction algorithms have been developed to process optical
interferometric data.  Using these algorithms may require substantial
expertise (for instance you may have to learn a specific programming
language). One of the objectives of the OPTICON JRA4 is to provide a common
graphical user interface (GUI) to drive such imaging algorithms in a
\emph{user friendly} way. For obvious reasons, we also want to avoid rewriting
the image reconstruction algorithms as much as possible.  A model of
interaction between these two pieces of software under these constraints has
to be developed. This document aims at specifying a common software interface
for image reconstruction algorithms.

On the image reconstruction side, this involves specifying exactly the inputs:
OI-FITS file for the data, the initial image, and other settings such as the
regularization and its hyper-parameters.\footnote{In a previous discussion,
John Young suggested to have a text file with a simple format to gives the
values of the input parameters. The format of this file must be simple and
such that it can be easily edited by a human and easily read by a software.
Here a different proposition is made which exploits the flexibility of the
FITS format.} Given these inputs, the image reconstruction algorithm should
start the reconstruction producing two kind of outputs:
\begin{description}
  \item[Intermediate outputs:]  These consist in iteration informations
  printed on the standard output (for instance)  and captured by the GUI and
  also the current solution in the form of an image, the corresponding model
  of the measurements (\emph{e.g.}, the complex visibilities).  These outputs
  will be used by the GUI to display the current solution, to compare the
  actual data with the current model, \emph{etc.}

  \item[Final outputs:]  These are essentially the final image and consistent
  information about the pixel size, the orientation, \emph{etc.} but also
  about the input data (\emph{i.e.} to avoid a \emph{traceability issue}), the
  reconstruction method and its parameters.
\end{description}

A few related points have to be considered:
\begin{enumerate}
  \item The command line image reconstruction algorithms should (or not?)
  switch off their own display capabilities.

  \item There must be a way to interrupt the image reconstruction algorithms
  so as to make sure they do release all their resources (no zombie process,
  undetached shared memory, etc.).

  \item Restarting an image reconstruction (to continue the iterations or
  perhaps with a few settings changed) should simply be a matter to re-start
  the algorithm with, as its input image, the last produced image.

  \item  In case of errors, we must adopt some kind of conventions to make
  sure that the GUI can properly account for the errors.
\end{enumerate}

%The GUI will be developed at JMMC, presumably in Java as it is mostly portable
%and they already use this language for their developments. Furthermore, they
%have a mastering for the kind of software wrapping we want to have. For
%instance, having a GUI which drives a command line program is exploited by
%LITpro, the model fitting software of the JMMC.


To summarize, on entry:
\begin{itemize}
\item the data;
\item image parameters (dimensions, pixel size, \emph{etc.});
\item initial image;
\item regularization and hyper-parameters;
\end{itemize}
On output (for the final and intermediate results):
\begin{itemize}
\item the current/final image;
\item the model of the data;
\item other relevant information;
\end{itemize}


To make the interfacing as simple as possible, it is proposed in this document
to use a single entry file with al input parameters and data and a single
output file with all the results.  The flexibility of the FITS file format is
exploited to store all these informations and data.

All image reconstruction software must be able to run from the command line
and with two arguments: the names of the input and output files (possibly with
some flag to indicate that the software must be run in a specific way).   For
instance:
\begin{verbatim}
  command --batch input output
\end{verbatim}


%==============================================================================
\section{Input format}

All existing algorithms take their input data in the OI-FITS format
\citep{Pauls_et_al-2005-oifits}.  This format stores the optical
interferometric data in FITS binary tables.  More generally, a FITS file
\citep{Pence_et_al-2010-FITS} consists in a number of so-called header data
units (HDU), each HDU having an header part with scalar parameters
specified by FITS keywords\footnote{A FITS keyword consist in upper case
  latin letters, digits, hyphen or underscore characters and has at least
  one character and at most height characters.} and a data part.  The
header part obeys a specific format but it is textual and easy to read for
an human.  The data part usually consist in binary data stored in various
forms.  In this document we only consider FITS \emph{images} (that is a
multidimensional array of values of the same data type) and FITS
\emph{binary tables}.  These are sufficient for our purposes.

By adding FITS HDU or introducing new FITS keywords, it is possible to provide
any additional data (\emph{e.g.} the initial image) and parameters required to
run the image reconstruction process.  Thus, it is proposed that the input
file be a valid OI-FITS file (resulting from the merging of all the optical
interferometric data to process) completed by HDU with additional binary data
needed for the reconstruction (for instance, the initial image) and with
scalar input settings provided as the values of FITS keywords.  This is
detailed in the following subsections.


\subsection{Scalar input parameters}

\renewcommand{\ROW}[3]{\KEYWORD{#1} & #2 &#3 \\}
\renewcommand{\ROWTITLE}[1]{\multicolumn{3}{c}{\textbf{#1}} \\
  Keyword & Type & Description \\}

\begin{table}
\caption{FITS keywords used to specify the input parameters.  The \emph{Image
Parameters} are given in the HDU which stores the initial image (identified by
the value of the \KEYWORD{INIT\_IMG} keyword), all other parameters are in the
primary HDU. \label{tab:input-params}}

\begin{tabular}{lcl}
  \hline
  \ROWTITLE{Data Selection}
  \hline
  \ROW{TARGET}{string}{Identifier of the target object to reconstruct}
  \ROW{WAVE\_MIN}{real}{Minimum wavelength to select (in meters)}
  \ROW{WAVE\_MAX}{real}{Maximum wavelength to select (in meters)}
  \ROW{USE\_VIS}{logical}{Use complex visibility data if any}
  \ROW{USE\_VIS2}{logical}{Use squared visibility data if any}
  \ROW{USE\_T3}{logical}{Use triple visibility data if any}
  \hline
  \ROWTITLE{Algorithm Settings}
  \hline
  \ROW{INIT\_IMG}{string}{Identifier of the initial image}
  \ROW{MAXITER}{integer}{Maximum number of iterations to run}
  \ROW{RGL\_NAME}{string}{Name of the regularization method}
  \ROW{RGL\_WGT}{real}{Weight of the regularization}
  \ROW{RGL\_ALPH}{real}{Parameter $\alpha$ of the regularization}
  \ROW{RGL\_BETA}{real}{Parameter $\beta$ of the regularization}
  \ROW{RGL\_PRIO}{string}{Identifier of the HDU with the prior image}
  \hline
  \ROWTITLE{Image Parameters}
  \hline
  \ROW{NAXIS1}{integer}{First dimension of the image}
  \ROW{NAXIS2}{integer}{Second dimension of the image}
  \ROW{CDELT1}{real}{Pixel size along first dimension of the image (in radians)}
  \ROW{CDELT2}{real}{Pixel size along second dimension of the image (in radians)}
  \hline
  \ROWTITLE{Algorithm Results}
  \hline
  \ROW{LAST\_IMG}{string}{Identifier of the final image}
  \hline
\end{tabular}
\end{table}

The first HDU, the so-called primary HDU, of a FITS file can only contains a,
possibly empty, FITS image. As optical interferometric data are all stored in
binary table, the primary HDU is almost empty. It is proposed that the input
(scalar) parameters be provided as FITS keywords in the primary HDU. There are
however some exceptions like the pixel size and the dimensions of the image
which are specified by those of the initial image (see below) which may be
stored in an other HDU.  Non-scalar parameters, for instance the prior image
for maximum entropy regularization, are stored in dedicated HDU (as images or
binary tables) and are identified by the value of their \KEYWORD{EXTNAME}
keyword.  Table~\ref{tab:input-params} gives some examples of the input
parameters.


\subsection{Data selection}

To keep things simple, any sophisticated selection, merging or editing of
the data should be done by a separate tool.  The image reconstruction
softwares shall assume that they receive as their inputs clean data.  There
are however a few parameters devoted to the selection of data.  The
\KEYWORD{TARGET} keyword is set with the identifier of the target object to
reconstruct.  The value of this keyword should match one of the identifiers
in the column \KEYWORD{TARGET\_ID} in OI-TARGET binary table of the OI-FITS
file.  In order to restrict the accounting of some type of interferometric
data, keywords \KEYWORD{USE\_VIS}, \KEYWORD{USE\_VIS2} and
\KEYWORD{USE\_T3} should be set with a logical value depending whether to
use complex visibility, the squared visibility, and the triple-product data
if any.  Not all algorithms can use all type of data and the value of these
keywords should be set (in the output file) so as to reflect what was
really used.

\oops{Shall we have \KEYWORD{USE\_VISA}/\KEYWORD{USE\_VISP} to distinguish
  amplitude/phase data? And likely, \KEYWORD{USE\_T3A}/\KEYWORD{USE\_T3P}
  for the triple-product data?}

Wavelength selection is discussed in the next section.

\oops{Open question: do the image reconstruction software have to honor the
  \KEYWORD{FLAG} column of the OI data?  Personnaly (Éric) I think that the
  answer is yes.  Then the meaning of this flag should be recalled.}


\subsection{Wavelength range}

The primary objective is to consider monochromatic image reconstruction.  The
interferometric data is generally available at many wavelengths.  The result
will, in fact, be a gray image of the target built from the data in the
wavelength range (inclusive) specified by the FITS keywords \KEYWORD{WAVE\_MIN}
and \KEYWORD{WAVE\_MAX}.  \oops{The wavelength range is given in meters.}


\subsection{Initial image}

All reconstruction algorithms are iterative and require an initial image to
start with. The initial image is provided as a FITS image in one of the HDU of
the input file. The primary HDU can be used to store the initial image in the
primary HDU (however \emph{cf.} the discussion about which image to store in
the primary HDU). The pixel size and the dimensions of the initial image
determine that of the sought image.  The dimensions are given by the FITS
keywords \KEYWORD{NAXIS1} and \KEYWORD{NAXIS2} while the pixel size is given
by the FITS keywords \KEYWORD{CDELT1} and \KEYWORD{CDELT2} (both values must
be the same).

It is not intended that the image reconstruction algorithm be able to deal
with any possible world coordinate system (WCS) nor with any possible
coordinate units.  The images are therefore stored according to a given WCS
with given units.  First image axis corresponds to the right ascension (RA)
and second image axis corresponds to the declination (DEC).  \oops{Specify
  the units, the orientation and the ordering of pixels?}  The same
conventions (WCS, units and orientation) hold for any output image produced
by the reconstruction algorithm.  For any external software to display
correctly the images, the parameters of the WCS must be completely and
correctly specified. \oops{This requires to set \KEYWORD{CRPIXn},
  \KEYWORD{CRVALn}, \KEYWORD{CTYPEn}, and \KEYWORD{CUNITn} keywords.  Other
  WCS keywords like \KEYWORD{CROTAn},
  \KEYWORD{PC\VARIABLE{i}\_\VARIABLE{j}},
  \KEYWORD{DC\VARIABLE{i}\_\VARIABLE{j}}, should not be specified as their
  default values (according to FITS standard) are suitable for us.}

\oops{Another issue: for most reconstruction software if not all, what is
  relevant is the relative coordinates not the absolute ones.  We may
  specify that the reference pixel given by keywords
  \KEYWORD{CRPIX\VARIABLE{i}} indicate where (in pixel units) is the center
  of the field of view (FOV) and that, by default, this is the geometrical
  center of the FOV.  Then the pixel size is given by the keywords
  \KEYWORD{CDELT\VARIABLE{i}} which must have the same absolute value (the
  sign can be used to cope with the orientation, \emph{explain how}).  The
  keywords \KEYWORD{CRVAL\VARIABLE{i}} can, optionally, be specified to
  indicate the absolute coordinates of the center of the FOV and to convert
  relative coordinates to absolute ones.}

%==============================================================================
\section{Output format}

It is proposed that the output format be as similar as possible as the input
format.  The output file must provide the reconstructed image but also some
informations for analyzing the result.  In particular, as each imaging
algorithm may implement its own method to estimate the complex visibilities
given the image of the object, it is necessary that the model of every fitted
data be computed by the algorithm itself rather than by another tool.


\subsection{Final/current image}

To compare the initial and the final images, they must be stored in different
HDU, the FITS keyword \KEYWORD{EXTNAME} is used to distinguish them.  As
previously explained, the dimensions, pixel size and orientation of the output
image(s) are the same as the initial image.

%The value of the FITS keyword \KEYWORD{EXTNAME} is \STRING{INITIAL IMAGE} and
%\STRING{FINAL IMAGE} for the initial and final images respectively.

As most image viewers\footnote{not all softwares deal with FITS files} are
only capable of displaying the image stored in the primary HDU, we suggest to
store the initial image in the primary HDU for the input file but to store the
final or current image in the primary HDU of the output file.  The idea is to
have the most relevant image stored in the primary HDU.  For the image
reconstruction algorithm and for any software designed to display or analyze
the results, the different images are distinguished by their names (given by
their \KEYWORD{EXTNAME} keyword).  \oops{OK but the last image should have a
specific name, perhaps set \KEYWORD{LAST\_IMG} keyword in the primary HDU with
name of the final image and let all reconstructed images be called as:
\begin{quote}
\KEYWORD{EXTNAME = 'OUTPUT\VARIABLE{n}'}
\end{quote}
with \VARIABLE{n} the iteration number.}

In order to continue the iterations of a previous reconstruction run, the
image reconstruction algorithm may be started with the final image instead of
the initial one.   To that end, there must be some means to specify which
image to start with for the image reconstruction software.  This is the
purpose of the \KEYWORD{INIT\_IMG} keyword (see Table~\ref{tab:input-params})
in the primary HDU which indicates the name of the initial image.



\subsection{Model of the data}

The OI-FITS format \citep{Pauls_et_al-2005-oifits} specifies that optical
interferometric data be stored in binary tables as columns with specific
names.  As there is no restriction that the tables only contain the columns
specified by the standard, we propose to store the model of the data in the
same tables by adding new columns.  The names of the new columns are listed in
Table~\ref{tab:model-columns}.  In this way it is very easy to compare the
actual data and their model as computed from the reconstructed image and the
instrumental model assumed by the reconstruction algorithm.  Another advantage
of this convention is that the same format can be exploited to store the
values given by model fitting software.

\renewcommand{\ROW}[2]{\texttt{#1} & \texttt{D(\textsl{NWAVE})} & #2 \\}
\renewcommand{\ROWTITLE}[1]{\multicolumn{3}{c}{\textbf{#1}}\\}

\begin{table}
\caption{Colums inserted in OI-FITS binary tables to store the values given by
the model.  \texttt{\textsl{NWAVE}} is the number of wavelengths.
\label{tab:model-columns}}
\begin{tabular}{lcl}
\hline
\hline
\ROWTITLE{New columns in \texttt{OI\_VIS} table}
Label & Format & Description \\
\hline
\ROW{VISAMPMODEL}{Model of the visibility amplitude}
\ROW{VISPHIMODEL}{Model of the visibility phase in degrees}
\hline
\hline
\ROWTITLE{New column in \texttt{OI\_VIS2} table}
Label & Format & Description \\
\hline
\ROW{VIS2MODEL}{Model of the squared visibility}
\hline
\hline
\ROWTITLE{New columns in \texttt{OI\_T3} table}
Label & Format & Description \\
\hline
\ROW{T3AMPMODEL}{Model of the triple-product amplitude}
\ROW{T3PHIMODEL}{Model of the triple-product phase in degrees}
\hline
\end{tabular}
\end{table}

%==============================================================================
\appendix
\section{Glossary}

\begin{description}
\item[DEC] Declination;
\item[FITS] Flexible Image Transport System;
\item[FOV] Field of view;
\item[GUI] Graphical User Interface;
\item[OI-FITS] Optical Interferometric exchange data format;
\item[RA] Right Ascension;
\item[WCS] World Coordinate System;
\end{description}

%==============================================================================
\bibliographystyle{plainnat}
\bibliography{OI-Interface}

\end{document}

